---
title: Occurrence, risk assessment and elimination of chemicals of emerging concerns
  in wasterwater treatment plants in Western Kenya
author: "Ali The Analyst"
date: "2023-03-20"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r cars}
library(readxl)
mydata<-read_excel("RR data (1).xlsx")
```

installing and loading  the required libraries

```{r pressure, echo=FALSE}
library(magrittr)
library(dplyr)
```
#Nzoia
Filtering Nzoia_EFF with highest detection of above 200
```{r}
Nzoia_EFF<-mydata %>% filter(Nzoia_EFF>=200) %>% select(WWTPs,Nzoia_EFF)
Nzoia_EFF
```

filtering Nzoia_INF with highest detection > 200 
```{r}
Nzoia<-mydata %>% filter (Nzoia_INF >= 200) %>% 
  select (Compound_class,WWTPs,Nzoia_INF,Nzoia_EFF)
Nzoia
```

Percentage removal of each chemical compound present in Nzoai with an influent value greater than 200
```{r}
per_removal<- ((Nzoia$Nzoia_INF - Nzoia$Nzoia_EFF)/ Nzoia$Nzoia_INF) * 100
per_removal_perchem<- data.frame(Nzoia$WWTPs,per_removal)
library(knitr)
knitr::kable(per_removal_perchem)
```

The Average concentration & STD concentration for each compound Class in Nzoia
```{r}
Nzoia_Summary<-Nzoia %>% group_by(Compound_class) %>% 
  summarise(average_INF = mean(Nzoia_INF), 
            std_INF = sd(Nzoia_INF),
            average_EFF = mean(Nzoia_EFF),
            std_EFF = sd(Nzoia_EFF))
knitr::kable(Nzoia_Summary)
```

Removal Efficiency for each compound class in Nzoia
```{r}
Nzoia_Efficiency<- Nzoia %>% group_by(Compound_class) %>%
  summarise(percentage_Efficiency = ((sum(Nzoia_INF)-sum(Nzoia_EFF))/sum(Nzoia_INF)) * 100)
knitr::kable(Nzoia_Efficiency)
```

#DLK
Filtering out the chemical compounds with influent values greater than 200
```{r}
DLK<-mydata %>% filter (DLK_INF >= 200) %>% select (Compound_class,WWTPs,DLK_INF,DLK_EFF)
DLK
```

Removal efficiency for each chemical componet in DLK
```{r}
percentage_removal<- ((DLK$DLK_INF - DLK$DLK_EFF)/ DLK$DLK_INF) * 100
per_removal_perchem<- data.frame(DLK,percentage_removal)
knitr::kable(per_removal_perchem)
```

 Average concentration & STD concentration for each compound Class in DLK
```{r}
DLK_Summary<-DLK %>% group_by(Compound_class) %>% 
  summarise(average_INF = mean(DLK_INF), 
            std_INF = sd(DLK_INF),
            average_EFF = mean(DLK_EFF),
            std_EFF = sd(DLK_EFF))
knitr::kable(DLK_Summary)
```
 
Removal efficiency for each compound class in DLK
```{r}
DLK_Efficiency<- DLK %>% group_by(Compound_class) %>%
  summarise(percentage_Efficiency = ((sum(DLK_INF)-sum(DLK_EFF))/sum(DLK_INF)) * 100)
knitr::kable(DLK_Efficiency)

```

#Moi
filtering the chemical substances with an influent value greater than 200
```{r}
Moi<-mydata %>% filter (Moi_INF >= 200) %>% select (Compound_class,WWTPs,Moi_INF,Moi_EFF)
Moi
```
removal efficiency for each chemical subtance in Moi
```{r}
percentage_removal<- ((Moi$Moi_INF - Moi$Moi_EFF)/ Moi$Moi_INF) * 100
per_removal_perchem<- data.frame(Moi$WWTPs,percentage_removal)
knitr::kable(per_removal_perchem)
```

Average concentration & STD concentration for every compound Class in Moi
```{r}
Moi_Summary<-Moi %>% group_by(Compound_class) %>% 
  summarise(average_INF = mean(Moi_INF), 
            std_INF = sd(Moi_INF),
            average_EFF = mean(Moi_EFF),
            std_EFF = sd(Moi_EFF))
knitr::kable(Moi_Summary)
```

Removal efficiency for each compound class in Moi
```{r}
Moi_Efficiency<- Moi %>% group_by(Compound_class) %>%
  summarise(percentage_Efficiency = ((sum(Moi_INF)-sum(Moi_EFF))/sum(Moi_INF)) * 100)
knitr::kable(Moi_Efficiency)
```

#Eldo
filtering the chemical substances with an influent value greater than 200
```{r}
Eldo<-mydata %>% filter (Eldo_INF >= 200) %>% select (Compound_class,WWTPs,Eldo_INF,Eldo_EFF)
```
Removal efficiency for every chemical substance  in Eldo
```{r}
percentage_removal<- ((Eldo$Eldo_INF - Eldo$Eldo_EFF)/ Eldo$Eldo_INF) * 100
per_removal_perchem<- data.frame(Eldo$WWTPs,percentage_removal)
knitr::kable(per_removal_perchem)
```

 Average concentration & STD concentration for each compound Class in Eldo
```{r}
Eldo_Summary<-Eldo %>% group_by(Compound_class) %>% 
  summarise(average_INF = mean(Eldo_INF), 
            std_INF = sd(Eldo_INF),
            average_EFF = mean(Eldo_EFF),
            std_EFF = sd(Eldo_EFF))
knitr::kable(Eldo_Summary)
```
 
 Removal efficiency for each compound class in DLK
```{r}
Eldo_Efficiency<- Eldo %>% group_by(Compound_class) %>%
  summarise(percentage_Efficiency = ((sum(Eldo_INF)-sum(Eldo_EFF))/sum(Eldo_INF)) * 100)
knitr::kable(Eldo_Efficiency)
```
 
